R Notebook

### 2022-05-30 code.Rmd
Update code.Rmd

house <- house %>% 
  mutate(Price = strtoi(str_replace_all(str_sub(Sold.Price, 2, -4), ",", ""))) %>%
  rename(nbhd = Neighborhood, bed = Bed.Rooms, fullBath = Full.Baths, halfBath = Half.Baths, style = Style, age = Year.Built) %>%
  rename(lotSize = Lot.Size, lastPrice = Last.Price, originalPrice = Original.Price, soldPrice = Price) %>%
  select(- Sold.Price) %>%
  rename(marketDays = Days.on.Market) %>%
  separate(col = Sold.Date, into = c("monthSold", "daySold", "yearSold"), sep = "/") %>%
  mutate(yearSold = strtoi(yearSold) + 2000) %>%
  mutate(daySold = strtoi(daySold), monthSold = strtoi(monthSold)) %>%
  mutate(age = yearSold - strtoi(age)) %>%
  select(- Property.Marketing.Period) %>%
  mutate(originalPrice = strtoi(str_replace_all(str_sub(originalPrice, 2, -4), ",", ""))) %>%
  mutate(lastPrice = strtoi(str_replace_all(str_sub(lastPrice, 2, -4), ",", ""))) %>%
  select(- X) %>%
  mutate(lotSize = as.double(lotSize)) %>%
  select(- Address) %>%
  mutate(marketDays = strtoi(marketDays))
```

```{r}
#some final tidying

@@

house <- house %>%
  select(- lotSize) %>%     #too many NAs
  select(- originalPrice) %>%
  select(- lastPrice) %>%  #we omit these two predictors because they are obviously way too similar to the response variable
  mutate(nbhd = replace(nbhd, nbhd == "Battelfield Area", "Battlefield Area"), nbhd = replace(nbhd, nbhd == "princeton Ridge", "Princeton Ridge"), nbhd = replace(nbhd, nbhd == "LIttlebrook", "Littlebrook"), nbhd = replace(nbhd, nbhd == "griggs Farm", "Griggs Farm"), nbhd = replace(nbhd, nbhd == "PrettyBrook Area", "Pretty Brook Area"), nbhd = replace(nbhd, nbhd == "Palmer Sq", "Palmer Square"), nbhd = replace(nbhd, nbhd == "Institute", "Institute Area"), nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"), nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"), nbhd = replace(nbhd, nbhd == "Carnegie Lake", "Carnegie Lake Area"), nbhd = replace(nbhd, nbhd == "Town", "Township")) %>%     #fix area names
  filter(! nbhd %in% c("Edgerstoune", "Northridge", "Rushbrook", "Russell Estates", "The Preserve", "Governors Lane", "Princeton Borough", "Preserve", "Queenston")) %>%  #delete areas with few houses
  filter(! style %in% badStyles$style) %>%  #delete styles with at most 3 houses
  mutate(style = replace(style, style == "Bi-level", "Bi-Level"), style = replace(style, style == "End Unit", "End-Unit"), style = replace(style, style == "Townhouse", "Townhome"))  #fix style names

house <- house %>%
  na.omit()
```

@@

set.seed(42)
houseSplit <- houseNorm
houseSplit$id <- sample(0:9, size = nrow(houseSplit), replace = TRUE)
houseSplit %>%
  group_by(id) %>%
  summarize(count = n())
```

```{r}
#define evaluation metrics: returns R-squared and RMS error

@@

#performance prints two numbers: the first one average R-squared, the second one average RMSE (standard deviation of the ten RMS)

@@

#performance(houseSplit, linearModel)
#performance(houseSplit, forwardModel)
#performance(houseSplit, backwardModel)

#R-Squared acceptable, but there is a large variance
```

```{r}
#we will use the glmnet package to build regularized regression models
#these include ridge, lasso, and elastic net

@@

#performance(houseSplit, ridgeModel)
#the ridge model has roughly the same RMSE, but slightly lower R2
#performance(houseSplit, lassoModel)
#lasso has slightly lower RMSE, and a slightly higher R2 compared to ridge, but still lower than the linear model
#performance(dummySplitWithResponse, elasticNetModel)
#elastic net is between ridge and lasso in terms of both R2 and variance
```

```{r}
#finally we try principal component regression (PCR)
#this is worth trying because many predictors are already correlated

@@

#performance(dummySplitWithResponse, PCRModel)
#This one has a higher R squared, but as a trade-off it also has high variance
```

@@

models <- data.frame(lm = performance(houseSplit, linearModel), forward = performance(houseSplit, forwardModel), backward = performance(houseSplit, backwardModel), ridge = performance(houseSplit, ridgeModel), lasso = performance(houseSplit, lassoModel), ela = performance(dummySplitWithResponse, elasticNetModel), pcr = performance(dummySplitWithResponse, PCRModel))

models
```

>>>>

house <- house %>% 
  mutate(Price = strtoi(str_replace_all(str_sub(Sold.Price, 2, -4), ",", ""))) %>%
  rename(nbhd = Neighborhood, 
         bed = Bed.Rooms, 
         fullBath = Full.Baths, 
         halfBath = Half.Baths, 
         style = Style, 
         age = Year.Built) %>%
  rename(lotSize = Lot.Size, 
         lastPrice = Last.Price, 
         originalPrice = Original.Price, 
         soldPrice = Price) %>%
  select(- Sold.Price) %>%
  rename(marketDays = Days.on.Market) %>%
  separate(col = Sold.Date, into = c("monthSold", "daySold", "yearSold"), sep = "/") %>%
  mutate(yearSold = strtoi(yearSold) + 2000) %>%
  mutate(daySold = strtoi(daySold), monthSold = strtoi(monthSold)) %>%
  mutate(age = yearSold - strtoi(age)) %>%
  select(- Property.Marketing.Period) %>%
  mutate(originalPrice = strtoi(str_replace_all(str_sub(originalPrice, 2, -4), ",", ""))) %>%
  mutate(lastPrice = strtoi(str_replace_all(str_sub(lastPrice, 2, -4), ",", ""))) %>%
  select(- X) %>%
  mutate(lotSize = as.double(lotSize)) %>%
  select(- Address) %>%
  mutate(marketDays = strtoi(marketDays))

#some final tidying

@@

house <- house %>%
  select(- lotSize) %>%     #too many NAs
  select(- originalPrice) %>%
  select(- lastPrice) %>%  #omit these two predictors because they are obviously way too similar to the response variable
  mutate(nbhd = replace(nbhd, nbhd == "Battelfield Area", "Battlefield Area"), 
         nbhd = replace(nbhd, nbhd == "princeton Ridge", "Princeton Ridge"), 
         nbhd = replace(nbhd, nbhd == "LIttlebrook", "Littlebrook"), 
         nbhd = replace(nbhd, nbhd == "griggs Farm", "Griggs Farm"), 
         nbhd = replace(nbhd, nbhd == "PrettyBrook Area", "Pretty Brook Area"), 
         nbhd = replace(nbhd, nbhd == "Palmer Sq", "Palmer Square"), 
         nbhd = replace(nbhd, nbhd == "Institute", "Institute Area"), 
         nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"), 
         nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"), 
         nbhd = replace(nbhd, nbhd == "Carnegie Lake", "Carnegie Lake Area"), 
         nbhd = replace(nbhd, nbhd == "Town", "Township")) %>%     #fix area names
  filter(! nbhd %in% c("Edgerstoune", "Northridge", "Rushbrook", "Russell Estates", 
                       "The Preserve", "Governors Lane", "Princeton Borough", 
                       "Preserve", "Queenston")) %>%  #delete areas with few houses
  filter(! style %in% badStyles$style) %>%  #delete styles with at most 3 houses
  mutate(style = replace(style, style == "Bi-level", "Bi-Level"), 
         style = replace(style, style == "End Unit", "End-Unit"), 
         style = replace(style, style == "Townhouse", "Townhome"))  #fix style names

house <- na.omit(house)
```

@@

set.seed(42)
houseSplit <- houseNorm
houseSplit$id <- sample(0:9, size = nrow(houseSplit), replace = TRUE)
# houseSplit %>%
#  group_by(id) %>%
#  summarize(count = n())

#define evaluation metrics: returns R-squared and RMS error.

@@

#performance prints two numbers: 
#the first one is the average R-squared across all 10 cross-validations
#the second one is the average RMSE (standard deviation of the ten RMS)

@@

#performance(houseSplit, linearModel)
#performance(houseSplit, forwardModel)
#performance(houseSplit, backwardModel)
#all three models behave fairly similarly: R-Squared is acceptable, but there is a large variance

#we will use the glmnet package to build regularized regression models
#these include ridge, lasso, and elastic net

@@

#performance(houseSplit, ridgeModel)
#the ridge model has roughly the same RMSE, but slightly lower R2
#performance(houseSplit, lassoModel)
#lasso has slightly lower RMSE, and a slightly higher R2 compared to ridge, but still lower than the linear model
#performance(dummySplitWithResponse, elasticNetModel)
#elastic net is between ridge and lasso in terms of both R2 and variance

#finally we try principal component regression (PCR)
#this is worth trying because many predictors are already correlated
#we can fine-tune the parameter ncomp, which has been set to 30 (half of the total number of predictors)

@@

#performance(dummySplitWithResponse, PCRModel)
#This one has a higher R squared, but as a trade-off it also has higher variance
```

@@

models <- data.frame(lm = performance(houseSplit, linearModel), 
                     forward = performance(houseSplit, forwardModel), 
                     backward = performance(houseSplit, backwardModel), 
                     ridge = performance(houseSplit, ridgeModel), 
                     lasso = performance(houseSplit, lassoModel), 
                     ela = performance(dummySplitWithResponse, elasticNetModel), 
                     pcr = performance(dummySplitWithResponse, PCRModel))
models
```
